### 2016-01-05 docs/proposal/proposal.tex
Add papers to read

\documentclass{article}

% GRAPHICS
\usepackage{graphicx}

% HYPERLINKS
\usepackage{hyperref}

% DOCUMENT PADDING AND MARGINS
\usepackage{titlesec}
\usepackage[margin=1.5in]{geometry}
\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}
\titlespacing*{\section}{0pt}{2ex}{0ex}
\titlespacing*{\subsection}{0pt}{2ex}{-2ex}
\titlespacing*{\subsubsection}{0pt}{2ex}{-2ex}

\begin{document}

% TITLE
\title{Autonomous Quadrotor Landing on a Moving Platform}
\author{Stan Brown \& Chris Choi}
\maketitle

\section*{Introduction}
Landing a quadrotor on a moving platform.

\section*{Previoius and Related Work}

\section*{Methods}

\section*{Milestones}
\begin{enumerate}
	\item{Identify hardware (camera, extra sensors...)}
	\item{Identify potential algorithmic solutions}
	\item{Implement solutions in softwaree}
	\item{Implement solutions in hardware}
	\item{Calibration of sensors}
	\item{Computer vision ID landing zone}
	\item{Motion Planning to land}
	\item{Perform the above successfully in simulation}
	\item{Perform the above successfully in real life}
\end{enumerate}

\section*{Potential Problems}
\begin{itemize}
	\item{Gazebo may have poor wind simulations}
	\item{Quadrotor ground effects (GP calcuting the error term?)}
\end{itemize}

\bibliography{proposal}{}
\bibliographystyle{plain}

\end{document}