### 2016-04-05 assignments/presentation/presentation.tex
Update carrot controller and start presentation deck

\documentclass{beamer}
\begin{document}
  \begin{frame}
    \frametitle{This is the first slide}
    %Content goes here
  \end{frame}
  \begin{frame}
    \frametitle{This is the second slide}
    \framesubtitle{A bit more information about this}
    %More content goes here
  \end{frame}
% etc
\end{document}