### 2016-04-05 assignments/presentation/presentation.tex
adding image caputure code

\documentclass{beamer}

\usepackage{media9}
\usetheme{Warsaw}

\begin{document}
\title{Quadrotor landing on a moving platform}
\subtitle{}
\author[]{Stan Brown \and Chris Choi}

\frame{\titlepage}

  \begin{frame}
    \frametitle{Introduction}
     \textbf{Presentation Outline}
     \begin{itemize}     
        \item Project Overview and Goals       
        \item Hardware used in the project
        \item System overview
        \item Approaches used to simplify the problem
        \item Milestones and Progress through the term       
    \end{itemize}
  \end{frame}

   \begin{frame}
    \frametitle{Project Overview and Goals}
     The goal of this work was to design and implement a set of methods that could be used 			to land a quadrotor on a moving platform. In our approach we broke the problem down 		in to a series of steps and goals along with the requirements for each step.
     \begin{itemize}     
        \item Develop a way to land on a stationary location on the ground
        \item Expand the methods employed in the first step to work on a moving platform 						operating in a planar, disturbance free environment
        \item Further expand the methods to be able to account for wind, non planar motion 					of the platform (ie identify and handle cases were the landing pad is 						located on a sloped surface).  
    \end{itemize}     
\end{frame}

 \begin{frame}
    \frametitle{Hardware}
     The hardware used in this project was selected from a set of parts that we had laying 			around the WAVELAB. 
     \begin{itemize}    
        \item Develop a way to land on a stationary location on the ground
        \item Expand the methods employed in the first step to work on a moving platform 						operating in a planar, disturbance free environment
        \item Further expand the methods to be able to account for wind, non planar motion 					of the platform (ie identify and handle cases were the landing pad is 						located on a sloped surface).  
    \end{itemize}     
\end{frame}

% etc
\end{document}